---
title: "Reproducible Research - Course Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Activity Monitoring

This project makes use of data from a personal activity monitoring device. The data was downloaded from  
[Activity Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip).

```{r, echo=TRUE}
library(dplyr)
fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileURL, "activity.zip")
unzip("activity.zip")
activities <- read.csv("activity.csv")

## Create a new column called newDate and convert the existing date column to 
## a valid date type
activities["newDate"] <- as.Date(activities$date,'%Y-%m-%d') 

## remove NA's
completeactivities <- activities[complete.cases(activities),]

## group by Date
activityGroups <- completeactivities %>% group_by(newDate)
stepCountByDate <- summarize(activityGroups, steps=sum(steps))
hist(stepCountByDate$steps,xlab = "Number of Steps", main = "Histogram of total number of steps taken each day")
```

## Mean and Median of Steps taken each day

```{r mean and median, echo=TRUE}
print(paste('Mean: ', mean(stepCountByDate$steps,na.rm = TRUE)))
print(paste('Median: ',median(stepCountByDate$steps, na.rm = TRUE)))
```

## Time Series plot of the average number of steps taken

```{r time series plot,echo = TRUE}
plot(stepCountByDate$newDate,stepCountByDate$steps, ylab = "Average Number of Steps", xlab = "Date", 
     type = 'l',main = 'Time Series of Average Number of Steps')
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r max steps, echo = TRUE}
completeactivities[completeactivities$steps == max(completeactivities$steps),]
```

## Code to describe and show a strategy for imputing missing data

Looking at the summary to analyze missing data

```{r summary of missing data, echo = TRUE}
summary(activities)
```


We have 2304 NAs in steps. Lets assume that 0 steps were taken at that interval.
And look at summary again
```{r replace missing values, echo = TRUE}
activities[is.na(activities)] <- 0
summary(activities)
```
## Histogram of the total number of steps taken each day after missing values are imputed
```{r histogram after replacing missing values, echo = TRUE}
hist(stepCountByDate$steps,xlab = "Number of Steps", main = "Histogram of total number of steps taken each day")
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r, echo=TRUE, fig.height=20,fig.width=40}
wdays<-c("Monday","Tuesday","Wednesday","Thursday","Friday")
wends<-c("Saturday","Sunday")

wdayActivities <- completeactivities[weekdays(completeactivities$newDate) %in% wdays,]
wendActivities <- completeactivities[weekdays(completeactivities$newDate) %in% wends,]

par(mfrow=c(2,2))
plot(wendActivities$interval,wendActivities$steps, xlab = 'Intervals', ylab = 'Steps', main = 'Activites during weekend', type = 'l')
plot(wdayActivities$interval,wdayActivities$steps, xlab = 'Intervals', ylab = 'Steps', main = 'Activites during weekday', type = 'l')

```


